---
title: "Homework 1"
author: "Emma Warshaw"
date: "2022-09-23"
output: github_document
---

```{r}
library(tidyverse)
```

# Problem 1

This problem will focus on use of inline R code, plotting and the behavior of `ggplot` for variables of different types.  

I will first download the data package `penguins` with a command directly in the console per the assignment instructions. 

Now, I'll load the dataset in the following code: 

```{r}
data("penguins", package = "palmerpenguins")
penguins
```

Let's calculate the mean flipper length for use in the discussion of the data set: 

```{r}
mean(penguins$flipper_length_mm, na.rm = TRUE)
```

Short description of the dataset: 
The 'penguins' dataset has `nrow=344` and `ncol=8`.  There are a total of eight variables including: species, island where the penguin lives, bill length (millimeters), bill depth (millimeters), flipper length (millimeters), body mass (grams), sex and year the data was recorded.  The average flipper length for penguins in the dataset is 200.92 millimeters.  

Now make a scatterplot of `flipper_length_mm` (y) vs `bill_length_mm` (x).  Color the points using the `species` variable (adding `color=...`) inside the `aes` in your `ggplot` code.  Then, export the scatterplot to the project directory using `ggsave`.   

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() 
ggsave("scatterplot_problem_1.pdf")
```

# Problem 2

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.  

Create a data frame comprised of: 
  - a random sample of size 10 from a standard Normal distribution
  - a logical vector indicating whether elements of the sample are greater than 0
  - a character vector of length 10 
  - a factor vector of length 10, with 3 different factor "levels"

```{r}
problem2_df = tibble(
  normal = rnorm(10),
  logical = normal > 0,
  character = character(length = 10),
  factor = factor(character(length = 10),(levels = 3))
)
```

Try to take the mean of each variable in the data frame.  What works and what doesn't? 

For the normal variable: 
```{r}
pull(problem2_df, var = normal)
mean(problem2_df$normal)
```
This works.  

For the logical variable:
```{r}
pull(problem2_df, var = logical)
mean(problem2_df$logical)
```
This also works.  

For the character variable:
```{r}
pull(problem2_df, var = character)
mean(problem2_df$character)
```
This does not work because the argument is not numeric or logical. 

For the factor variable: 
```{r}
pull(problem2_df, var = factor)
mean(problem2_df$factor)
```
This also does not work because the argument is not numeric or logical. 



